---
title: "Homework 2"
author: "Yan Song"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

First, define a path to the dataset. 

```{r}
path_to_data = "./data/Trash-Wheel-Collection-Totals-7-2020-1.xlsx"
```


Read the Mr. Trashwheel dataset. 

```{r}
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

Read precipitation data! For 2018 and 2017.

```{r}
precip_2018 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-7-2020-1.xlsx",
		sheet = "2018 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)

precip_2017 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-7-2020-1.xlsx",
		sheet = "2017 Precipitation",
		skip = 1
	) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
```

Now combine annual precipitation dataframes. In the following code chunk, I create a "helper" tibble that contains pairs of numeric and character ways of representing month, and then merge that (using month number as a key) with the precipitation dataset. This technique is one I use often when I need to recode a moderate or large number of values for a variable. 

```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)

precip_df = 
	bind_rows(precip_2018, precip_2017)

precip_df =
	left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


## Problem 2
Read and clean NYC Transit data.

```{r}
nyc_transit_df = read.csv ("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", na.strings = c("","NA")) %>% 
janitor::clean_names() %>% 
select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
mutate(entry = recode(entry, 'YES' = TRUE, 'NO' = FALSE), 
       entry = as.logical(entry))

```
Description

The NYC transit data contains information related to each entrance and exit for each subway station in NYC.
The resulting dataset contains information on line, station name, station latitude, station logitude, routes served, entry, vending, entrance type and ADA compliance.And the entry variable is converted from character to logical variable. The dimension of the resulting data is `r nrow(nyc_transit_df)` x `r ncol(nyc_transit_df)`. 

```{r}
# Question 1
station_num = distinct(nyc_transit_df, line, station_name)

# Question 2
ada_num = distinct(nyc_transit_df, line, station_name,ada) %>% filter(ada == TRUE)

# Question 3
no_vending_num = nyc_transit_df %>%
  filter(vending == 'NO', entry == TRUE)
entry_num = nyc_transit_df %>%
  filter(vending == 'NO')

no_vending_prop = nrow(no_vending_num)/nrow(entry_num)*100
```
* There are `r count(station_num)` distinct stations.
* There are `r count(ada_num)` stations are ADA compliant.
* The proportion of station entrance/ exits without vending allow entrance is `r round(no_vending_prop, digits = 2)`%. 

```{r}
nyc_transit_tidy = nyc_transit_df %>%
  mutate (route8 = as.character(route8),
         route9 = as.character(route9),
         route10 = as.character(route10),
         route11 = as.character(route11)) %>%
  pivot_longer (
    route1:route11,
    names_to = "route_name",
    values_to = "route_num")
```
* There are `r count (distinct(nyc_transit_tidy, line, station_name, route_num) %>% filter(route_num == 'A'))` distinct stations serve the A train. Of the stations that serve the A train, there are `r count(distinct(nyc_transit_tidy, line, station_name, route_num, ada) %>% filter(route_num == 'A', ada == TRUE))` ADA compliant.

## Problem 3 

First, we read and clean the data in pols-month.csv.
```{r}

pols_month = read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    month = month.abb[month],
    president = ifelse(prez_gop == 1, "republican", "democratic")
    ) %>% 
  select(-prez_gop, -prez_dem, -day) %>% 
  arrange(year, month) 

```

Second, we read and clean the data in snp.csv.
```{r}

snp = read_csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"),sep = "/") %>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    month = month.abb[month]
  ) %>% 
  select(year, month, everything(),-day) 
  
```

Third, we read and tidy the umemployment data.
```{r}

unemployment = read_csv(file = "./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(
    "jan":"dec",
    names_to = "month",
    values_to = "unemployment_percentage"
  ) %>% 
  arrange(year, month) 

```

Then, we combine the above three datasets together. 
```{r}
pols = 
  left_join(pols_month, snp) 

pols_tidy = 
  left_join(pols, unemployment)

```

Description

* There are `r nrow(pols)` rows and `r ncol(pols)` columns in the pols_month dataset. The key variable which contains row keys is year and month in this dataset. The range of years of the pols_dataset is (`r range(pull(pols,year))`).

* There are `r nrow(snp)` rows and `r ncol(snp)` columns in the snp dataset. The key variable which contains row keys is year and month in this dataset. The range of years of the pols_dataset is (`r range(pull(snp,year))`).


* There are `r nrow(unemployment)` rows and `r ncol(unemployment)` columns in the employment data. The key variable which contains row keys is year and month in this dataset. The range of years of the pols_dataset is (`r range(pull(unemployment,year))`).
